---
title: "Breakdown of pangenome partition by species and COG category"
author: "Gavin Douglas"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    theme: cerulean
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(ComplexHeatmap))
suppressPackageStartupMessages(library(circlize))
suppressPackageStartupMessages(library(gridtext))
```

# Preprocess

## Read in COG annot information
```{r read_in}
intact_annot <- readRDS("/data1/gdouglas/projects/accessory_vs_pseudogene/summary/intact_and_both_cluster_annot.rds")
intact_annot <- intact_annot[-which(intact_annot$COG_category == "-"), ]
intact_annot <- intact_annot[-which(intact_annot$COG_category == ""), ]

# For intact clusters, filter out clusters outside size range (at least based on what clusters they're in).
cluster_filt_lengths <- read.table(file = "/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_filt_lengths_and_additional.tsv.gz",
                                   header = TRUE, sep = "\t", row.names = 1)

intact_annot <- intact_annot[which(rownames(intact_annot) %in% rownames(cluster_filt_lengths)), ]

pseudo_annot <- readRDS("/data1/gdouglas/projects/accessory_vs_pseudogene/summary/intergenic_pseudo_cluster_COG_annot.rds")
pseudo_annot <- pseudo_annot$pseudo
pseudo_annot <- pseudo_annot[-which(pseudo_annot$majority_COG_category == ""), ]

rownames(pseudo_annot) <- pseudo_annot$cluster
pseudo_annot <- pseudo_annot[, c(2, 3)]
colnames(pseudo_annot) <- c("COG", "COG_category")

intact_annot <- intact_annot[, c("all_COG", "COG_category")]
colnames(intact_annot) <- c("COG", "COG_category")

COG_category_annot <- rbind(intact_annot, pseudo_annot)

cluster_pangenome_categories <- readRDS("/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_pangenome_categories.rds")

# Read in COG category db info.
COG_category_descrip <- read.table("/data1/gdouglas/db/COG_definitions/COG_category_descrip.tsv",
                                   header = FALSE, sep="\t", stringsAsFactors = FALSE, row.names = 1)

# Ignore several COG categories (eukaryote-associated primarily)
COG_category_descrip <- COG_category_descrip[which(! rownames(COG_category_descrip) %in% c("A", "B", "Y", "Z")), , drop = FALSE]

```

# Heatmap-prep function

```{r prep_function}
prep_heatmap_input <- function(clusters_to_include) {

  # Subset to specific clusters only.
  cluster_pangenome_categories_subset <- cluster_pangenome_categories
  for (sp in names(cluster_pangenome_categories_subset)) {

   for (type in names(cluster_pangenome_categories_subset[[sp]])) {
    
      for (partition in names(cluster_pangenome_categories_subset[[sp]][[type]])) {
       
        cluster_pangenome_categories_subset[[sp]][[type]][[partition]] <- intersect(cluster_pangenome_categories_subset[[sp]][[type]][[partition]],
                                                                                    clusters_to_include)
         
      }
     
     }

  }
  
  all_species <- names(cluster_pangenome_categories_subset)
  
  pangenome_category_breakdown <- data.frame(matrix(NA, nrow = length(all_species), ncol = 12))
  rownames(pangenome_category_breakdown) <- all_species
  colnames(pangenome_category_breakdown) <- c("intact_ultra.cloud", "intact_other.cloud", "intact_shell", "intact_soft.core",
                                              "both_ultra.cloud", "both_other.cloud", "both_shell", "both_soft.core",
                                              "pseudo_ultra.cloud", "pseudo_other.cloud", "pseudo_shell", "pseudo_soft.core")
  
  for (sp in all_species) {
    
    for (type in c("intact", "both", "pseudo")) {
      
      for (partition in c("ultra.cloud", "other.cloud", "shell", "soft.core")) {
          column_name <- paste(type, partition, sep = "_")
          pangenome_category_breakdown[sp, column_name] <- length(cluster_pangenome_categories_subset[[sp]][[type]][[partition]])
      }

    }

  }

  pangenome_category_breakdown_rel.by.category <- pangenome_category_breakdown
  pangenome_category_breakdown_rel.by.pangenome <- pangenome_category_breakdown
  
  pseudo_colnames <- paste("pseudo", c("ultra.cloud", "other.cloud", "shell", "soft.core"), sep = "_")
  intact_colnames <- paste("intact", c("ultra.cloud", "other.cloud", "shell", "soft.core"), sep = "_")
  mixed_colnames <- paste("both", c("ultra.cloud", "other.cloud", "shell", "soft.core"), sep = "_")
  
  ultra.cloud_colnames <- paste(c("intact", "both", "pseudo"), "ultra.cloud", sep = "_")
  other.cloud_colnames <- paste(c("intact", "both", "pseudo"), "other.cloud", sep = "_")
  shell_colnames <- paste(c("intact", "both", "pseudo"), "shell", sep = "_")
  soft.core_colnames <- paste(c("intact", "both", "pseudo"), "soft.core", sep = "_")
  
  for (sp in all_species) {
    
    sp_total_pseudo <- sum(pangenome_category_breakdown[sp, pseudo_colnames])
    sp_total_intact <- sum(pangenome_category_breakdown[sp, intact_colnames])
    sp_total_mixed <- sum(pangenome_category_breakdown[sp, mixed_colnames])
    
    sp_total_ultra.cloud <- sum(pangenome_category_breakdown[sp, ultra.cloud_colnames])
    sp_total_other.cloud <- sum(pangenome_category_breakdown[sp, other.cloud_colnames])
    sp_total_shell <- sum(pangenome_category_breakdown[sp, shell_colnames])
    sp_total_soft.core <- sum(pangenome_category_breakdown[sp, soft.core_colnames])

    pangenome_category_breakdown_rel.by.category[sp, pseudo_colnames] <- (pangenome_category_breakdown_rel.by.category[sp, pseudo_colnames] / sp_total_pseudo) * 100
    pangenome_category_breakdown_rel.by.category[sp, intact_colnames] <- (pangenome_category_breakdown_rel.by.category[sp, intact_colnames] / sp_total_intact) * 100
    pangenome_category_breakdown_rel.by.category[sp, mixed_colnames] <- (pangenome_category_breakdown_rel.by.category[sp, mixed_colnames] / sp_total_mixed) * 100
  
    pangenome_category_breakdown_rel.by.pangenome[sp, ultra.cloud_colnames] <- (pangenome_category_breakdown_rel.by.pangenome[sp, ultra.cloud_colnames] / sp_total_ultra.cloud) * 100
    pangenome_category_breakdown_rel.by.pangenome[sp, other.cloud_colnames] <- (pangenome_category_breakdown_rel.by.pangenome[sp, other.cloud_colnames] / sp_total_other.cloud) * 100
    pangenome_category_breakdown_rel.by.pangenome[sp, shell_colnames] <- (pangenome_category_breakdown_rel.by.pangenome[sp, shell_colnames] / sp_total_shell) * 100
    pangenome_category_breakdown_rel.by.pangenome[sp, soft.core_colnames] <- (pangenome_category_breakdown_rel.by.pangenome[sp, soft.core_colnames] / sp_total_soft.core) * 100
    
  }
  
  pangenome_category_breakdown_rel.by.category <- round(pangenome_category_breakdown_rel.by.category, 1)
  pangenome_category_breakdown_rel.by.pangenome <- round(pangenome_category_breakdown_rel.by.pangenome, 1)
  pangenome_category_breakdown_rel.by.pangenome <- pangenome_category_breakdown_rel.by.pangenome[, c(ultra.cloud_colnames, other.cloud_colnames, shell_colnames, soft.core_colnames)]
  
  pangenome_category_breakdown_by.category <- pangenome_category_breakdown[, colnames(pangenome_category_breakdown_rel.by.category)]
  pangenome_category_breakdown_by.pangenome <- pangenome_category_breakdown[, colnames(pangenome_category_breakdown_rel.by.pangenome)]
  
  # Get text to show in heatmaps
  pangenome_category_breakdown_rel.by.category_char <- apply(pangenome_category_breakdown_rel.by.category, 2, function(x) { paste(x, "%\n", sep = "") })
  pangenome_category_breakdown_by.category_char <- apply(pangenome_category_breakdown_by.category, 2, function(x) { paste("(", x, ")", sep = "") })
  
  for (i in 1:ncol(pangenome_category_breakdown_rel.by.category_char)) {
    pangenome_category_breakdown_rel.by.category_char[, i] <- paste(pangenome_category_breakdown_rel.by.category_char[, i],
                                                                    pangenome_category_breakdown_by.category_char[, i], sep = "")
  }
  
  pangenome_category_breakdown_rel.by.pangenome_char <- apply(pangenome_category_breakdown_rel.by.pangenome, 2, function(x) { paste(x, "%\n", sep = "") })
  pangenome_category_breakdown_by.pangenome_char <- apply(pangenome_category_breakdown_by.pangenome, 2, function(x) { paste("(", x, ")", sep = "") })
  
  for (i in 1:ncol(pangenome_category_breakdown_rel.by.pangenome_char)) {
    pangenome_category_breakdown_rel.by.pangenome_char[, i] <- paste(pangenome_category_breakdown_rel.by.pangenome_char[, i],
                                                                    pangenome_category_breakdown_by.pangenome_char[, i], sep = "")
  }
  
  return(list(by_category = pangenome_category_breakdown_rel.by.category,
              by_category_fill = pangenome_category_breakdown_rel.by.category_char,
              by_pangenome = pangenome_category_breakdown_rel.by.pangenome,
              by_pangenome_fill = pangenome_category_breakdown_rel.by.pangenome_char))

}
```

# Analysis

## Overall breakdown heatmaps {.tabset}

```{r overall_prep_tables}
all_clusters <- as.character()

for (sp in names(cluster_pangenome_categories)) {

   for (type in names(cluster_pangenome_categories[[sp]])) {
    
      for (partition in names(cluster_pangenome_categories[[sp]][[type]])) {
       
        all_clusters <- c(all_clusters,
                          cluster_pangenome_categories[[sp]][[type]][[partition]])
         
      }
     
    }

}

all_clusters <- unique(all_clusters)

overall_prepped_files <- prep_heatmap_input(all_clusters)

# Save tables as files for future reference.
write.table(x = overall_prepped_files$by_category,
            file = "/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_breakdown_tables/all.clusters_percent_by_type.tsv",
            col.names = NA, row.names = TRUE, sep = "\t", quote = FALSE)

by_category_fill_to_write <- gsub("\n", " ", overall_prepped_files$by_category_fill)
rownames(by_category_fill_to_write) <- rownames(overall_prepped_files$by_category)
write.table(x = by_category_fill_to_write,
            file = "/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_breakdown_tables/all.clusters_percent_by_type_fill.tsv",
            col.names = NA, row.names = TRUE, sep = "\t", quote = FALSE)

write.table(x = overall_prepped_files$by_pangenome,
            file = "/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_breakdown_tables/all.clusters_percent_by_partition.tsv",
            col.names = NA, row.names = TRUE, sep = "\t", quote = FALSE)

by_pangenome_fill_to_write <- gsub("\n", " ", overall_prepped_files$by_pangenome_fill)
rownames(by_pangenome_fill_to_write) <- rownames(overall_prepped_files$by_pangenome)
write.table(x = by_pangenome_fill_to_write,
            file = "/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_breakdown_tables/all.clusters_percent_by_partition_fill.tsv",
            col.names = NA, row.names = TRUE, sep = "\t", quote = FALSE)
```


### By cluster type (intact, mixed, or pseudogene)

```{r overall_heatmap_by_intact_all, fig.width = 10, fig.height = 6}
draw(column_title = "Cluster breakdown by cluster type, for all clusters (including those lacking COG category annotation)",
     Heatmap(matrix = as.matrix(overall_prepped_files$by_category),
             name = "Percent",
             row_names_side = "left",
             row_labels = gt_render(paste("*", gsub("_", " ", rownames(overall_prepped_files$by_category)), "*", sep = "")),
             column_labels = c("Ultra-cloud", "Other-cloud", "Shell", "Soft-core", "Ultra-cloud", "Other-cloud", "Shell", "Soft-core", "Ultra-cloud", "Other-cloud", "Shell", "Soft-core"),
             column_split = c("Intact", "Intact", "Intact", "Intact", "Mixed", "Mixed", "Mixed", "Mixed", "Pseudogene", "Pseudogene", "Pseudogene", "Pseudogene"),
             cluster_rows = FALSE,
             column_gap = unit(5, "mm"),
             cluster_columns = FALSE,
             column_names_rot = 45,
             col = colorRamp2(c(0, 50, 100), c("slateblue1","white", "red")),
             cell_fun = function(j, i, x, y, width, height, fill) {
               grid.text(overall_prepped_files$by_category_fill[i, j], x, y, gp = gpar(fontsize = 10, fontface="bold"))
             }))
```


### By pseudogene partition

```{r overall_heatmap_by_partition_all, fig.width = 10, fig.height = 6}
draw(column_title = "Cluster breakdown by pangenome partition, for all clusters (including unannotated)",
     Heatmap(matrix = as.matrix(overall_prepped_files$by_pangenome),
            name = "Percent",
            column_labels = c("Intact", "Mixed", "Pseudogene", "Intact", "Mixed", "Pseudogene", "Intact", "Mixed", "Pseudogene", "Intact", "Mixed", "Pseudogene"),
            row_names_side = "left",
            row_labels = gt_render(paste("*", gsub("_", " ", rownames(overall_prepped_files$by_pangenome)), "*", sep = "")),
            column_split = c("a) Ultra-cloud", "a) Ultra-cloud", "a) Ultra-cloud", "b) Other-cloud", "b) Other-cloud", "b) Other-cloud", "c) Shell", "c) Shell", "c) Shell", "d) Soft-core", "d) Soft-core", "d) Soft-core"),
            cluster_rows = FALSE,
            column_gap = unit(5, "mm"),
            cluster_columns = FALSE,
            column_names_rot = 45,
            col = colorRamp2(c(0, 50, 100), c("slateblue1","white", "red")),
            cell_fun = function(j, i, x, y, width, height, fill) {
              grid.text(overall_prepped_files$by_pangenome_fill[i, j], x, y, gp = gpar(fontsize = 10, fontface="bold"))
            })
)
```


## Many pseudogenes have no COG annotation (especially those in soft-core partition)

```{r unannot_heatmap, fig.width = 10, fig.height = 6}
unannot_clusters <- all_clusters[which(! all_clusters %in% rownames(COG_category_annot))]

all_species <- names(cluster_pangenome_categories)
  
pangenome_category_breakdown_unannot <- data.frame(matrix(NA, nrow = length(all_species), ncol = 12))
rownames(pangenome_category_breakdown_unannot) <- all_species
colnames(pangenome_category_breakdown_unannot) <- c("intact_ultra.cloud", "intact_other.cloud", "intact_shell", "intact_soft.core",
                                                    "both_ultra.cloud", "both_other.cloud", "both_shell", "both_soft.core",
                                                    "pseudo_ultra.cloud", "pseudo_other.cloud", "pseudo_shell", "pseudo_soft.core")

pangenome_category_breakdown_unannot_percent <- pangenome_category_breakdown_unannot

for (sp in all_species) {
  
  for (type in c("intact", "both", "pseudo")) {
    
    for (partition in c("ultra.cloud", "other.cloud", "shell", "soft.core")) {
      column_name <- paste(type, partition, sep = "_")
      pangenome_category_breakdown_unannot[sp, column_name] <- length(which(cluster_pangenome_categories[[sp]][[type]][[partition]] %in% unannot_clusters))
      pangenome_category_breakdown_unannot_percent[sp, column_name] <- (pangenome_category_breakdown_unannot[sp, column_name] /
                                                                        length(cluster_pangenome_categories[[sp]][[type]][[partition]])) * 100
    }
    
  }
  
}

pangenome_category_breakdown_unannot_percent_char <- apply(pangenome_category_breakdown_unannot_percent, 2, function(x) { sprintf('%.1f', x) })
pangenome_category_breakdown_unannot_char <- apply(pangenome_category_breakdown_unannot, 2, as.character)
pangenome_category_breakdown_unannot_heatmap_fill <- pangenome_category_breakdown_unannot_percent_char
for (i in 1:nrow(pangenome_category_breakdown_unannot_heatmap_fill)) {
  for(j in 1:ncol(pangenome_category_breakdown_unannot_heatmap_fill)) {
    pangenome_category_breakdown_unannot_heatmap_fill[i, j] <- paste(pangenome_category_breakdown_unannot_percent_char[i, j],
                                                                     "%\n(",
                                                                     pangenome_category_breakdown_unannot_char[i, j],
                                                                     ")",
                                                                     sep = "")
  }
}

draw(column_title = "Percent of clusters without COG annotation (out of total for that cluster type, species, and pangenome partition)",
     Heatmap(matrix = as.matrix(pangenome_category_breakdown_unannot_percent),
        name = "Percent",
        row_names_side = "left",
        row_labels = gt_render(paste("*", gsub("_", " ", rownames(pangenome_category_breakdown_unannot_percent)), "*", sep = "")),
        column_labels = c("Ultra-cloud", "Other-cloud", "Shell", "Soft-core", "Ultra-cloud", "Other-cloud", "Shell", "Soft-core", "Ultra-cloud", "Other-cloud", "Shell", "Soft-core"),
        column_split = c("Intact", "Intact", "Intact", "Intact", "Mixed", "Mixed", "Mixed", "Mixed", "Pseudogene", "Pseudogene", "Pseudogene", "Pseudogene"),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        column_names_rot = 45,
        column_gap = unit(5, "mm"),
        col = colorRamp2(c(0, 50, 100), c("slateblue1","white", "red")),
        cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(pangenome_category_breakdown_unannot_heatmap_fill[i, j], x, y, gp = gpar(fontsize = 10, fontface="bold"))
        })
)


# Save these tables too:
write.table(x = pangenome_category_breakdown_unannot_percent,
            file = "/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_breakdown_tables/unannot_percent_by_table.cell.tsv",
            col.names = NA, row.names = TRUE, sep = "\t", quote = FALSE)

pangenome_category_breakdown_unannot_heatmap_fill_to_write <- gsub("\n", " ", pangenome_category_breakdown_unannot_heatmap_fill)
rownames(pangenome_category_breakdown_unannot_heatmap_fill_to_write) <- rownames(pangenome_category_breakdown_unannot_percent)
write.table(x = pangenome_category_breakdown_unannot_heatmap_fill_to_write,
            file = "/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_breakdown_tables/unannot_percent_by_table.cell_fill.tsv",
            col.names = NA, row.names = TRUE, sep = "\t", quote = FALSE)
```


## Clusters annotated with COG category {.tabset}

```{r annot_prep_tables}
annot_prepped_files <- prep_heatmap_input(rownames(COG_category_annot))

# Save tables as files for future reference.
write.table(x = annot_prepped_files$by_category,
            file = "/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_breakdown_tables/cog.category.annot_percent_by_type.tsv",
            col.names = NA, row.names = TRUE, sep = "\t", quote = FALSE)

annot.only_by_category_fill_to_write <- gsub("\n", " ", annot_prepped_files$by_category_fill)
rownames(annot.only_by_category_fill_to_write) <- rownames(annot_prepped_files$by_category)
write.table(x = annot.only_by_category_fill_to_write,
            file = "/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_breakdown_tables/cog.category.annot_percent_by_type_fill.tsv",
            col.names = NA, row.names = TRUE, sep = "\t", quote = FALSE)

write.table(x = annot_prepped_files$by_pangenome,
            file = "/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_breakdown_tables/cog.category.annot_percent_by_partition.tsv",
            col.names = NA, row.names = TRUE, sep = "\t", quote = FALSE)

annot.only_by_pangenome_fill_to_write <- gsub("\n", " ", annot_prepped_files$by_pangenome_fill)
rownames(annot.only_by_pangenome_fill_to_write) <- rownames(annot_prepped_files$by_pangenome)
write.table(x = annot.only_by_pangenome_fill_to_write,
            file = "/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_breakdown_tables/cog.category.annot_percent_by_partition_fill.tsv",
            col.names = NA, row.names = TRUE, sep = "\t", quote = FALSE)
```

### By cluster type (intact, mixed, or pseudogene)

```{r overall_heatmap_by_intact, fig.width = 10, fig.height = 6}
draw(column_title = "Cluster breakdown by cluster type, for COG-category annotated clusters",
     Heatmap(matrix = as.matrix(annot_prepped_files$by_category),
             name = "Percent",
             row_names_side = "left",
             row_labels = gt_render(paste("*", gsub("_", " ", rownames(annot_prepped_files$by_category)), "*", sep = "")),
             column_labels = c("Ultra-cloud", "Other-cloud", "Shell", "Soft-core", "Ultra-cloud", "Other-cloud", "Shell", "Soft-core", "Ultra-cloud", "Other-cloud", "Shell", "Soft-core"),
             column_split = c("Intact", "Intact", "Intact", "Intact", "Mixed", "Mixed", "Mixed", "Mixed", "Pseudogene", "Pseudogene", "Pseudogene", "Pseudogene"),
             cluster_rows = FALSE,
             column_gap = unit(5, "mm"),
             cluster_columns = FALSE,
             column_names_rot = 45,
             col = colorRamp2(c(0, 50, 100), c("slateblue1","white", "red")),
             cell_fun = function(j, i, x, y, width, height, fill) {
               grid.text(annot_prepped_files$by_category_fill[i, j], x, y, gp = gpar(fontsize = 10, fontface="bold"))
             }))
```


### By pseudogene partition

```{r overall_heatmap_by_partition, fig.width = 10, fig.height = 6}
draw(column_title = "Cluster breakdown by pangenome partition, for COG-category annotated clusters",
     Heatmap(matrix = as.matrix(annot_prepped_files$by_pangenome),
            name = "Percent",
            column_labels = c("Intact", "Mixed", "Pseudogene", "Intact", "Mixed", "Pseudogene", "Intact", "Mixed", "Pseudogene", "Intact", "Mixed", "Pseudogene"),
            row_names_side = "left",
            row_labels = gt_render(paste("*", gsub("_", " ", rownames(annot_prepped_files$by_pangenome)), "*", sep = "")),
            column_split = c("a) Ultra-cloud", "a) Ultra-cloud", "a) Ultra-cloud", "b) Other-cloud", "b) Other-cloud", "b) Other-cloud", "c) Shell", "c) Shell", "c) Shell", "d) Soft-core", "d) Soft-core", "d) Soft-core"),
            cluster_rows = FALSE,
            column_gap = unit(5, "mm"),
            cluster_columns = FALSE,
            column_names_rot = 45,
            col = colorRamp2(c(0, 50, 100), c("slateblue1","white", "red")),
            cell_fun = function(j, i, x, y, width, height, fill) {
              grid.text(annot_prepped_files$by_pangenome_fill[i, j], x, y, gp = gpar(fontsize = 10, fontface="bold"))
            })
)
```


## Breakdown by COG category {.tabset}

```{r COG_looping, results = 'asis', echo = FALSE, fig.width = 10, fig.height = 6}
for (COG_category in rownames(COG_category_descrip)) {
  
  cat('###', COG_category, "-", COG_category_descrip[COG_category, 1], '{.tabset}', '\n')
  
  COG_clusters <- rownames(COG_category_annot)[grep(COG_category, COG_category_annot$COG_category)]
  
  COG_prepped_files <- prep_heatmap_input(COG_clusters)
 
  
  write.table(x = COG_prepped_files$by_category,
            file = paste("/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_breakdown_tables/by_cog_category/", COG_category, ".annot_percent_by_type.tsv", sep = ""),
            col.names = NA, row.names = TRUE, sep = "\t", quote = FALSE)

  COG.category_by_category_fill_to_write <- gsub("\n", " ", COG_prepped_files$by_category_fill)
  rownames(COG.category_by_category_fill_to_write) <- rownames(COG_prepped_files$by_category)
  write.table(x = COG.category_by_category_fill_to_write,
              file = paste("/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_breakdown_tables/by_cog_category/", COG_category, ".annot_percent_by_type_fill.tsv", sep = ""),
              col.names = NA, row.names = TRUE, sep = "\t", quote = FALSE)

  write.table(x = COG_prepped_files$by_pangenome,
            file = paste("/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_breakdown_tables/by_cog_category/", COG_category, ".annot_percent_by_partition.tsv", sep = ""),
            col.names = NA, row.names = TRUE, sep = "\t", quote = FALSE)

  COG.category_by_pangenome_fill_to_write <- gsub("\n", " ", COG_prepped_files$by_pangenome_fill)
  rownames(COG.category_by_pangenome_fill_to_write) <- rownames(COG_prepped_files$by_pangenome)
  write.table(x = COG.category_by_pangenome_fill_to_write,
              file = paste("/data1/gdouglas/projects/accessory_vs_pseudogene/summary/cluster_breakdown_tables/by_cog_category/", COG_category, ".annot_percent_by_partition_fill.tsv", sep = ""),
              col.names = NA, row.names = TRUE, sep = "\t", quote = FALSE)
  
  
  cat('####', 'By cluster type', '\n')

  cat('\n\n')
  
  draw(column_title = paste("Pangenome partitions of intact vs pseudogene, for ", COG_category, "-annotated clusters", sep = ""),
       Heatmap(matrix = as.matrix(COG_prepped_files$by_category),
             name = "Percent",
             row_names_side = "left",
             row_labels = gt_render(paste("*", gsub("_", " ", rownames(COG_prepped_files$by_category)), "*", sep = "")),
             column_labels = c("Ultra-cloud", "Other-cloud", "Shell", "Soft-core", "Ultra-cloud", "Other-cloud", "Shell", "Soft-core", "Ultra-cloud", "Other-cloud", "Shell", "Soft-core"),
             column_split = c("Intact", "Intact", "Intact", "Intact", "Mixed", "Mixed", "Mixed", "Mixed", "Pseudogene", "Pseudogene", "Pseudogene", "Pseudogene"),
             cluster_rows = FALSE,
             column_gap = unit(5, "mm"),
             cluster_columns = FALSE,
             column_names_rot = 45,
             col = colorRamp2(c(0, 50, 100), c("slateblue1","white", "red")),
             cell_fun = function(j, i, x, y, width, height, fill) {
               grid.text(COG_prepped_files$by_category_fill[i, j], x, y, gp = gpar(fontsize = 10, fontface="bold"))
             })
  )
  
  cat('\n\n')
  
  cat('####', 'By pseudogene partition', '\n')
  
  draw(column_title = paste("Breakdown of intact vs cloud per pangenome partition, COG category:", COG_category),
       Heatmap(matrix = as.matrix(COG_prepped_files$by_pangenome),
            name = "Percent",
            column_labels = c("Intact", "Mixed", "Pseudogene", "Intact", "Mixed", "Pseudogene", "Intact", "Mixed", "Pseudogene", "Intact", "Mixed", "Pseudogene"),
            row_names_side = "left",
            row_labels = gt_render(paste("*", gsub("_", " ", rownames(COG_prepped_files$by_pangenome)), "*", sep = "")),
            column_split = c("a) Ultra-cloud", "a) Ultra-cloud", "a) Ultra-cloud", "b) Other-cloud", "b) Other-cloud", "b) Other-cloud", "c) Shell", "c) Shell", "c) Shell", "d) Soft-core", "d) Soft-core", "d) Soft-core"),
            cluster_rows = FALSE,
            column_gap = unit(5, "mm"),
            cluster_columns = FALSE,
            column_names_rot = 45,
            col = colorRamp2(c(0, 50, 100), c("slateblue1","white", "red")),
            cell_fun = function(j, i, x, y, width, height, fill) {
              grid.text(COG_prepped_files$by_pangenome_fill[i, j], x, y, gp = gpar(fontsize = 10, fontface="bold"))
            })
  )
  
  cat('\n\n')
   
}
```


# Session info {.tabset}

## Hide session info

## Show session info

Session details printed out for reproducibility.

```{r session_info}
sessionInfo()
```
